---
title: "Meta Analysis Practice"
author: "Paloma Cartwright"
date: "09/09/2021"
output: 
  html_document:
    toc: true 
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(metafor)
library(janitor)
library(tidyverse)
```
Reading in the data set from Gardner et. al. cleaned to contain only primates. 

```{r}
DatRich <- read_csv(here::here("Gardner_et_al_data_onlyprimates.csv")) 
```
## Calculating the Unstandardized Values

This is problematic because there is no way of knowing if the values and variables are the same type and unit

#### Calculate the Unstandardized Mean Difference 
Subtract the MeanvalueatLowFauna from the MeanvalueatHighFauna to get the mean difference

```{r}
DatRich <- DatRich %>% 
  mutate(u_mean_diff = MeanvalueatHighFauna - MeanvalueatLowFauna)

```

#### Calculate the pooled standard deviation 

```{r}
DatRich <- DatRich %>% 
  mutate(s_pooled = 
           sqrt((((HighFaunaN -1) * SDatHighFauna ^ 2) + ((LowFaunaN - 1) * SDatLowFauna ^ 2))
                /((HighFaunaN -1) + (LowFaunaN - 1))))

```

#### Calculate the mean difference standard error between high and low fauna

```{r}
DatRich <- DatRich %>% 
  mutate(umd_se = s_pooled * sqrt((1 / HighFaunaN) + (1 / LowFaunaN)))
```


## Calculating the Standardized Values 

#### Calculating the standardized mean difference

```{r}
DatRich <- DatRich %>% 
  mutate(smd = u_mean_diff / s_pooled)
```

#### Calculating the correction factor 

```{r}
DatRich <- DatRich %>% 
  mutate(j = (1 - (3 / (4 * (HighFaunaN + LowFaunaN - 2) - 1))))
```

#### Calculating the Hedge's g value 

```{r}
DatRich <- DatRich %>% 
  mutate(g = smd * j )
```

#### Calculating the standard error of Hedge's g 

When the sample sizes are small Cohen's d values can be biased so you correct for that by calculating the Hedge's g value

```{r}
DatRich <- DatRich %>% 
  mutate(hedgesg_se = sqrt(((HighFaunaN + LowFaunaN) / (HighFaunaN * LowFaunaN))
                           + ((g ^ 2)/(2 * (HighFaunaN + LowFaunaN)))))
```

## Fixed Effects Model 

The pooled effect is a weighted average. The assumption is that there is one true effect and any variation from that is due to sampling error. This is not necessarily a good assumption, so beware. But, if it is only sampling area that cause deviation from the true effects, then you can weight each study by their inverse variance. 

#### Calculate the inverse variance weighting 

```{r}
DatRich <- DatRich %>% 
  mutate( w = 1 / (hedgesg_se ^ 2))

```

#### Calculate the Theta 

```{r}
pooled_fixed_effect <- DatRich %>% 
  summarize(pooled_fixed_effect = (sum(g * w ))/ sum(w)) %>% 
  pull() # this is a dplyr function that just pulls the value out instead of fucking up the dataframe. It can also pull a whole vector 
```

## Using metafor 

Metafor can do everything that we just did by hand. So now that we know what we were doing, metafor is available for us 

`escalc()` - effect size calculation 

### Creating the Forest Plot 

```{r, fig.height= 7, fig.width=5}

DatRich <- escalc(n1i = LowFaunaN, n2i = HighFaunaN, m1i = MeanvalueatLowFauna, 
                  m2i = MeanvalueatHighFauna, sd1i = (SDatLowFauna), sd2i = (SDatHighFauna), 
                  data = DatRich, measure = "SMD", append = TRUE)
# SMD is Hedge's g in metafor 

Rich_Overall <- rma(yi, vi, method = "FE", data = DatRich)
forest(Rich_Overall) 

```

### Test for Publication Bias 

```{r, fig.height=5, fig.width=3}
par(mfrow = c(2, 1))
funnel(Rich_Overall, main="Standard Error")
funnel(Rich_Overall, yaxis="seinv", main="Inverse Standard Error")

```



